Increasing Productivity Amid Stable Nutrient Regimes in Rhode Island Lakes and Reservoirs

### 2019-05-24 manuscript/manuscript.Rmd
added URI WW numbers... Wow!

The analysis approach we used here, site-specific z-scores, is not a novel method, but using it to examine water quality trends is novel as we could find few examples of using it specifically for water quality trends [NEED TO DO THIS LIT SEARCH].  This approach, and in particular non-scaled anomalies, does have a long history in the analysis of trends in climate [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface].  We simply built on this methods and adapted it for use with long-term water quality trends.  And while other methods are valid and robust (e.g. [@Oliver_2017]), we chose averaged site-specific z-scores as they, in our opinion, provide readily interpretable results, especially for communicating to less technical audiences.  In addition, the site-specific z-scores are robust to variations in sampling effort and in the timing of inclusion of given sampling locations (i.e. added later in a time period or removed).  Lastly, this analysis is only possible because of the availability of sound, long-term data on water quality in Rhode Island.  In other words, without the URIWW data and the commitment and participation of nearly XXX volunteers our analysis would have been impossible.  Going forward, it is important to appreciate the role that volunteer monitoring and citizen science programs can play in the analysis of long term environmental trends.

>>>>

The analysis approach we used here, site-specific z-scores, is not a novel method, but using it to examine water quality trends is novel as we could find few examples of using it specifically for water quality trends [NEED TO DO THIS LIT SEARCH].  This approach, and in particular non-scaled anomalies, does have a long history in the analysis of trends in climate [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface].  We simply built on this methods and adapted it for use with long-term water quality trends.  And while other methods are valid and robust (e.g. [@Oliver_2017]), we chose averaged site-specific z-scores as they, in our opinion, provide readily interpretable results, especially for communicating to less technical audiences.  In addition, the site-specific z-scores are robust to variations in sampling effort and in the timing of inclusion of given sampling locations (i.e. added later in a time period or removed).  Lastly, this analysis is only possible because of the availability of sound, long-term data on water quality in Rhode Island.  In other words, without the URIWW data and the commitment and participation of more than 2500 volunteers over the years, our analysis would have been impossible.  Going forward, it is important to appreciate the role that volunteer monitoring and citizen science programs can play in the analysis of long term environmental trends.
